---
title: "Alternatives"
output:
  html_document:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(sf)
library(tidyverse)
library(data.table)
library(flextable)
library(ggspatial)

# Load shapefiles
shp_wards <- read_sf('./shapefiles/Wards_polygon.shp', query = "SELECT NAME FROM WARDS_polygon") # Calderdale ward shapefile
shp_newconst <- read_sf('./shapefiles/proposed_const.shp', 
                        query = "SELECT Constituen AS Constituency, Electorate FROM proposed_const WHERE Constituen IN (\'Halifax\', \'Calder Valley\', \'Batley and Hipperholme\', 'Huddersfield')") # proposed constituencies

# Load ward data
wards <- fread('./data/calderdale_wards.csv', select=c('Ward', 'ONS Code', 'const_2019', 'const_2023', 'abbreviation')) # 

# Load electorate from Electoral Commission ward data
bc_wards <- read_sf('./shapefiles/existing_and_prospective_wards_for_2023_review.shp', query = "SELECT ons_code, electorate AS Electorate, popdensity FROM existing_and_prospective_wards_for_2023_review") %>%
  as.data.frame() %>%
  subset(select = -c(geometry))  

# Join the data together
wards <- inner_join(bc_wards, wards, by=c('ons_code' = 'ONS Code')) %>%
  as.data.frame()

# Add ONS codes and electorate to wards shapefile
shp_wards <- full_join(shp_wards, wards, by = c("NAME" = "Ward"))

# Make it a shapefile again
shp_wards <- st_sf(shp_wards, sf_column_name = 'geometry')

# Compute centroids, so we can add labels
shp_wards <- cbind(shp_wards, st_coordinates(st_centroid(shp_wards)))

# Import polling districts and extract two-letter abbreviation
shp_polling_districts <- read_sf('./shapefiles/poll_dists.shp')
n_last <- 2
shp_polling_districts$LETTERS <- substr(shp_polling_districts$Name, nchar(shp_polling_districts$Name) - n_last + 1, nchar(shp_polling_districts$Name))

# Compute centroids
shp_polling_districts <- cbind(shp_polling_districts, st_coordinates(st_centroid(shp_polling_districts)))

# Import Polling stations
shp_polling_stations <- read_sf('./shapefiles/polling_districts.geojson')

# Define constituency colours
pal_colours <- c(
  "Calder Valley" = "Blue",
  "Halifax" = "Red",
  "Batley and Hipperholme" = "Grey",
  "Huddersfield" = "Grey"
  )

pal_shapes <- c(
  "Polling district" = 18
)

# Define plot theme
mytheme <- theme( 
  text = element_text(family = "Source Sans Pro"),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  plot.title = element_text(size = 18, face = "bold"),
  plot.subtitle = element_text(size = 12),
  axis.title = element_blank(),
  axis.line = element_blank(),
  axis.text = element_blank(),
  axis.ticks = element_blank(),
  legend.key = element_blank()
)

# Nudge labels slightly
shp_wards$nudge_y <- 0
shp_wards$nudge_x <- 0
shp_wards$nudge_y[shp_wards$abbreviation == "SB"] <- 200
shp_wards$nudge_x[shp_wards$abbreviation == "GS"] <- 650
shp_wards$nudge_x[shp_wards$abbreviation == "TOW"] <- 250
shp_wards$nudge_y[shp_wards$abbreviation == "BRI"] <- 200
```

This page sets out initial ideas for alternative proposals.

## Move Brighouse ward to Huddersfield

This proposal involves moving **Brighouse** ward to the **Huddersfield** constituency to the south. This would require that **Hipperholme and Lightcliffe** become part of **Halifax**, so as not to split **Calder Valley** in half. To equalise the numbers, **Sowerby Bridge** is moved to **Calder Valley**.

```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, out.width="100%"}
shp_wards_alt_1 <- shp_wards
shp_wards_alt_1$const_2023[shp_wards_alt_1$NAME == "Brighouse"] <- "Huddersfield"
shp_wards_alt_1$const_2023[shp_wards_alt_1$NAME == "Hipperholme And Lightcliffe"] <- "Halifax"
shp_wards_alt_1$const_2023[shp_wards_alt_1$NAME == "Sowerby Bridge"] <- "Calder Valley"
shp_wards_alt_1$const_2023[shp_wards_alt_1$NAME == "Warley"] <- "Halifax"
shp_wards_alt_1$const_2023[shp_wards_alt_1$NAME == "Rastrick"] <- "Calder Valley"

plt <- ggplot(shp_wards_alt_1)+
  mytheme+ 
#  geom_sf(data = shp_newconst, col = "grey75", alpha = 0, linetype = "dashed")+
  geom_sf(aes(fill = const_2023, alpha = popdensity), col = "grey20")+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )+
  scale_fill_manual(
      values = pal_colours,
      limits = names(pal_colours)[c(1:2,4)]
  )+
  scale_alpha(range = c(0.75, 1), guide = "none")+
  geom_sf(data = shp_polling_stations, aes(shape = "Polling district"), size = 1, fill = "grey20", col = "grey20")+ # Plot polling stations on map
  scale_shape_manual(values = pal_shapes, breaks = names(pal_shapes))+
  geom_label(
    data = shp_wards, 
    size = 2,
    aes(X, Y, label = abbreviation),
    nudge_x = shp_wards$nudge_x,
    nudge_y = shp_wards$nudge_y
    )+
  guides(colour = guide_legend(order = 2), 
         shape = guide_legend(order = 1))
plt
ggsave(
  './docs/downloads/alt_brig_hudds.svg',
  plot = plt
)
```
Download map as [SVG](./downloads/alt_brig_hudds.svg)
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, out.width="100%"}
wards_alt_1 <- wards
wards_alt_1$const_2023[wards_alt_1$Ward == "Brighouse"] <- "Huddersfield"
wards_alt_1$const_2023[wards_alt_1$Ward == "Hipperholme And Lightcliffe"] <- "Halifax"
wards_alt_1$const_2023[wards_alt_1$Ward == "Sowerby Bridge"] <- "Calder Valley"
wards_alt_1$const_2023[wards_alt_1$Ward == "Warley"] <- "Halifax"
wards_alt_1$const_2023[wards_alt_1$Ward == "Rastrick"] <- "Calder Valley"

t1 <- wards_alt_1 %>%
  select(const_2023, Electorate) %>%
  filter(const_2023 %in% c("Halifax", "Calder Valley")) %>%
  group_by(const_2023) %>%
  summarise(Electorate = sum(Electorate))
knitr::kable(t1, col.names = c("2023 Constituency", "Electorate"), format.args = list(big.mark = ","))
```

This would leave **Huddersfield** as a very large constituency, well outside the permitted range However, it may be possible to move some or all of **Almondbury** to a rebalanced Dewsbury constituency to leave **Huddersfield** within the permitted range.

## Move Rastrick ward to Huddersfield

Moving **Rastrick** to **Huddersfield** has much the same effect as moving **Brighouse**, except that it does not divide **Calder Valley**. This means the rest of the wards in Calderdale are unchanged.

```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, out.width="100%"}
shp_wards_alt_2 <- shp_wards
shp_wards_alt_2$const_2023[shp_wards_alt_1$NAME == "Rastrick"] <- "Huddersfield"
shp_wards_alt_2$const_2023[shp_wards_alt_1$NAME == "Hipperholme And Lightcliffe"] <- "Calder Valley"
shp_wards_alt_2$const_2023[shp_wards_alt_1$NAME == "Brighouse"] <- "Calder Valley"
shp_wards_alt_2$const_2023[shp_wards_alt_1$NAME == "Sowerby Bridge"] <- "Halifax"
shp_wards_alt_2$const_2023[shp_wards_alt_1$NAME == "Warley"] <- "Halifax"

plt <- ggplot(shp_wards_alt_2)+
  mytheme+ 
#  geom_sf(data = shp_newconst, col = "grey75", alpha = 0, linetype = "dashed")+
  geom_sf(aes(fill = const_2023, alpha = popdensity), col = "grey20")+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )+
  scale_fill_manual(
      values = pal_colours,
      limits = names(pal_colours)[c(1:2,4)]
  )+
  scale_alpha(range = c(0.75, 1), guide = "none")+
  geom_sf(data = shp_polling_stations, aes(shape = "Polling district"), size = 1, fill = "grey20", col = "grey20")+ # Plot polling stations on map
  scale_shape_manual(values = pal_shapes, breaks = names(pal_shapes))+
  geom_label(
    data = shp_wards, 
    size = 2,
    aes(X, Y, label = abbreviation),
    nudge_x = shp_wards$nudge_x,
    nudge_y = shp_wards$nudge_y
    )+
  guides(colour = guide_legend(order = 2), 
         shape = guide_legend(order = 1))
plt
ggsave(
  './docs/downloads/alt_rast_hudds.svg',
  plot = plt
)
```
Download map as [SVG](./downloads/alt_rast_hudds.svg)
```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, out.width="100%"}
wards_alt_2 <- wards
wards_alt_2$const_2023[shp_wards_alt_1$Ward == "Rastrick"] <- "Huddersfield"
wards_alt_2$const_2023[shp_wards_alt_1$Ward == "Hipperholme And Lightcliffe"] <- "Calder Valley"
wards_alt_2$const_2023[shp_wards_alt_1$Ward == "Brighouse"] <- "Calder Valley"
wards_alt_2$const_2023[shp_wards_alt_1$Ward == "Sowerby Bridge"] <- "Halifax"
wards_alt_2$const_2023[shp_wards_alt_1$Ward == "Warley"] <- "Halifax"


t1 <- wards_alt_2 %>%
  select(const_2023, Electorate) %>%
  filter(const_2023 %in% c("Halifax", "Calder Valley")) %>%
  group_by(const_2023) %>%
  summarise(Electorate = sum(Electorate))
knitr::kable(t1, col.names = c("2023 Constituency", "Electorate"), format.args = list(big.mark = ","))
```

## Divide Luddenden ward

This proposal involves dividing **Luddenden** ward between **Halifax** and **Calder Valley** constituencies. To bring **Calder Valley** into the permitted range, approximately 1,800 electors need to be moved.

This map shows the effect of moving the **Wainstalls (GA)**, **Luddenden (GB)**, **Luddendenfoot (GC)** and **Warley Wood (GD)** districts (shown in pink) into **Halifax**. This moves 3,374 electors and results in two constituencies of very similar size.

```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, out.width="100%"}
plt <- ggplot(shp_wards)+
  mytheme+ 
  geom_sf(aes(fill = const_2019, alpha = popdensity), col = "grey20")+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )+
  scale_fill_manual(
    values = pal_colours,
    limits = names(pal_colours)[1:2], # We don't need Batley and Hipperholme for this map
  )+
  scale_alpha(range = c(0.75, 1), guide = "none")+
  geom_sf(data = shp_polling_districts %>% filter(LETTERS %in% c("GA", "GB", "GC", "GD")),
          fill = "Pink",
          col = "grey20")+
  geom_sf(data = shp_polling_stations, aes(shape = "Polling district"), size = 1, fill = "grey20", col = "grey20")+ # Plot polling stations on map
  scale_shape_manual(values = pal_shapes, breaks = names(pal_shapes))+
  geom_label(
    data = shp_wards, 
    size = 2,
    aes(X, Y, label = abbreviation),
    nudge_x = shp_wards$nudge_x,
    nudge_y = shp_wards$nudge_y
  )
plt
ggsave(
  './docs/downloads/alt_ludd.svg',
  plot = plt
)
```
Download map as [SVG](./downloads/alt_ludd.svg)
``` {r, echo=FALSE}
constit <- c('Calder Valley', 'Halifax')
elect <- c(75450, 75100)
knitr::kable(data.frame(constit, elect), col.names = c("2023 Constituency", "Electorate"), format.args = list(big.mark = ","))
```

## Divide Hipperholme & Lightcliffe

The proposal involves dividing **Hipperholme and Lightcliffe** between **Halifax** and **Calder Valley** constituencies.

This map shows the effect of bringing the **Leeds Road (EA)** and **Hipperholme (EB)** polling districts into **Halifax**. 3,637 electors are moved, resulting in two neard-identical sized constituencies.

```{r, echo=FALSE, warning=FALSE, message=FALSE, error=FALSE, out.width="100%"}
plt <- ggplot(shp_wards)+
  mytheme+ 
  geom_sf(aes(fill = const_2019, alpha = popdensity), col = "grey20")+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )+
  scale_fill_manual(
    values = pal_colours,
    limits = names(pal_colours)[1:2], # We don't need Batley and Hipperholme for this map
  )+
  scale_alpha(range = c(0.75, 1), guide = "none")+
  geom_sf(data = shp_polling_districts %>% filter(LETTERS %in% c("EA", "EB")),
          fill = "Pink",
          col = "grey20")+
  geom_sf(data = shp_polling_stations, aes(shape = "Polling district"), size = 1, fill = "grey20", col = "grey20")+ # Plot polling stations on map
  scale_shape_manual(values = pal_shapes, breaks = names(pal_shapes))+
  geom_label(
    data = shp_wards, 
    size = 2,
    aes(X, Y, label = abbreviation),
    nudge_x = shp_wards$nudge_x,
    nudge_y = shp_wards$nudge_y
  )
plt
ggsave(
  './docs/downloads/alt_hl.svg',
  plot = plt
)
```

Download map as [SVG](./downloads/alt_hl.svg)

``` {r, echo=FALSE}
constit <- c('Calder Valley', 'Halifax')
elect <- c(75187, 75363)
knitr::kable(data.frame(constit, elect), col.names = c("2023 Constituency", "Electorate"), format.args = list(big.mark = ","))
```
The Electoral Commission's own report already includes proposals to divide wards in other constituencies:

> ... We propose that the Manor Castle ward be transferred from the existing Sheffield Central constituency into our proposed Sheffield Heeley constituency, bringing the former within the permitted electorate range. We then propose retaining the existing division of the Richmond ward between the Sheffield Heeley and Sheffield South East constituencies. We propose including the south-west part of the ward, including the areas of Four Lane Ends and Intake, in the Sheffield Heeley constituency, and the remainder of the ward, centred on Richmond, in the Sheffield South East constituency. We recognise that this divides a ward between constituencies, but consider that doing so enables the retention of a significant number of existing constituencies largely unchanged.
--- [Boundary Commission Initial proposals for Yorkshire and the Humber](https://boundarycommissionforengland.independent.gov.uk/wp-content/uploads/2021/06/2021-06-08-Initial-Proposals-Yorkshire-and-the-Humber-Region.pdf) ยง35

> ... We propose that the Dalton ward be divided, with the area around Kirkheaton being included in the proposed Dewsbury constituency, and the rest of the ward, centred on the Rawthorpe area, remaining in the Huddersfield constituency. We consider that not dividing a ward here would inevitably result in more disruption to local ties across the area.
--- ibid ยง61

> ... [W]e also propose that the Gipton & Harehills ward be divided between the Leeds Central and Leeds East constituencies, with the area in the west of the ward, centred on Harehills, being included in the Leeds Central constituency, and the remainder of the ward, centred on Gipton, being included in our proposed Leeds East constituency ... If we were not to divide a ward in the City of Leeds, we consider that there would inevitably be more disruption to local ties.
--- ibid ยง69

This would leave the proposed **Batley and Hipperholme** constituency with 66,845 electors - 2,879 short of the minimum required.