---
title: "Boundary Commission Proposal"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(sf)
library(tidyverse)
library(data.table)
library(flextable)
library(ggspatial)

# Load shapefiles
shp_wards <- read_sf('./shapefiles/Wards_polygon.shp', query = "SELECT NAME FROM WARDS_polygon") # Calderdale ward shapefile
shp_newconst <- read_sf('./shapefiles/proposed_const.shp', 
                        query = "SELECT Constituen AS Constituency, Electorate FROM proposed_const WHERE Constituen IN (\'Halifax\', \'Calder Valley\', \'Batley and Hipperholme\')") # proposed constituencies

# Load ward data
wards <- fread('./data/calderdale_wards.csv', select=c('Ward', 'ONS Code', 'const_2019', 'const_2023', 'abbreviation')) # 

# Load electorate from Electoral Commission ward data
bc_wards <- read_sf('./shapefiles/existing_and_prospective_wards_for_2023_review.shp', query = "SELECT ons_code, electorate AS Electorate, popdensity FROM existing_and_prospective_wards_for_2023_review") %>%
  as.data.frame() %>%
  subset(select = -c(geometry))  

# Join the data together
wards <- inner_join(bc_wards, wards, by=c('ons_code' = 'ONS Code')) %>%
  as.data.frame()

# Add ONS codes and electorate to wards shapefile
shp_wards <- full_join(shp_wards, wards, by = c("NAME" = "Ward"))

# Make it a shapefile again
shp_wards <- st_sf(shp_wards, sf_column_name = 'geometry')

# Compute centroids, so we can add labels
shp_wards <- cbind(shp_wards, st_coordinates(st_centroid(shp_wards)))

# Import Polling stations
shp_polling_stations <- read_sf('./shapefiles/polling_districts.geojson')

# Define constituency colours
pal_colours <- c(
  "Calder Valley" = "Blue",
  "Halifax" = "Red",
  "Batley and Hipperholme" = "Grey"
  )

pal_shapes <- c(
  "Polling district" = 18
)

# Define plot theme
mytheme <- theme( 
  text = element_text(family = "Source Sans Pro"),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  plot.title = element_text(size = 18, face = "bold"),
  plot.subtitle = element_text(size = 12),
  axis.title = element_blank(),
  axis.line = element_blank(),
  axis.text = element_blank(),
  axis.ticks = element_blank(),
  legend.key = element_blank()
)

# Nudge labels slightly
shp_wards$nudge_y <- 0
shp_wards$nudge_x <- 0
shp_wards$nudge_y[shp_wards$abbreviation == "SB"] <- 200
shp_wards$nudge_x[shp_wards$abbreviation == "GS"] <- 650
shp_wards$nudge_x[shp_wards$abbreviation == "TOW"] <- 250
shp_wards$nudge_y[shp_wards$abbreviation == "BRI"] <- 200
```

> We therefore propose that the Hipperholme and Lightcliffe ward be transferred from the Calder Valley constituency to a proposed Batley and Hipperholme constituency, based largely on the existing Batley and Spen constituency. In order to bring the two Calderdale constituencies within the permitted electorate range, the two wards of Sowerby Bridge and Warley, and the two wards of Brighouse and Rastrick are exchanged between the proposed Halifax and Calder Valley constituencies.

--- [Boundary Commission Initial proposals](https://boundarycommissionforengland.independent.gov.uk/wp-content/uploads/2021/06/2021-06-08-Initial-Proposals-Yorkshire-and-the-Humber-Region.pdf)

```{r, echo=FALSE, out.width="100%"}
ggplot(shp_wards)+
  mytheme+ 
  geom_sf(data = shp_newconst, col = "grey75", alpha = 0, linetype = "dashed")+
  geom_sf(aes(fill = const_2023, alpha = popdensity), col = "grey20")+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )+
  scale_fill_manual(
      values = pal_colours,
      limits = names(pal_colours)
  )+
  scale_alpha(range = c(0.75, 1), guide = "none")+
  geom_sf(data = shp_polling_stations, aes(shape = "Polling district"), size = 1, fill = "grey20", col = "grey20")+ # Plot polling stations on map
  scale_shape_manual(values = pal_shapes, breaks = names(pal_shapes))+
  geom_label(
    data = shp_wards, 
    size = 2,
    aes(X, Y, label = abbreviation),
    nudge_x = shp_wards$nudge_x,
    nudge_y = shp_wards$nudge_y
    )+
  guides(colour = guide_legend(order = 2), 
         shape = guide_legend(order = 1))
```

```{r, echo=FALSE}
t1 <- wards %>%
  select(const_2023, Electorate) %>%
  filter(const_2023 %in% c("Halifax", "Calder Valley")) %>%
  group_by(const_2023) %>%
  summarise(Electorate = sum(Electorate))
knitr::kable(t1, col.names = c("2023 Constituency", "Electorate"), format.args = list(big.mark = ","))
```

## Observations

* This achieves the commission's objectives -- both constituencies are of equal size and within the permitted range
* Five wards, made up of 43,586 electors, will move to a different constituency
  + This represents 29% of electors in the borough
* The integrity of Halifax constituency is largely maintained
* **Hipperholme and Lightcliffe** is now orphaned from the rest of the borough
  + Their new constituency will be split across multiple Local Authorities
  
Next: [Alternatives](./alternatives.html)