---
title: "Current Position"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(sf)
library(tidyverse)
library(data.table)
library(flextable)
library(ggspatial)

# Load shapefiles
shp_wards <- read_sf('./shapefiles/Wards_polygon.shp', query = "SELECT NAME FROM WARDS_polygon") # Calderdale ward shapefile
shp_newconst <- read_sf('./shapefiles/proposed_const.shp', 
                        query = "SELECT Constituen AS Constituency, Electorate FROM proposed_const WHERE Constituen IN (\'Halifax\', \'Calder Valley\', \'Batley and Hipperholme\')") # proposed constituencies

# Load ward data
wards <- fread('./data/calderdale_wards.csv', select=c('Ward', 'ONS Code', 'const_2019', 'const_2023', 'abbreviation')) # 

# Load electorate from Electoral Commission ward data
bc_wards <- read_sf('./shapefiles/existing_and_prospective_wards_for_2023_review.shp', query = "SELECT ons_code, electorate AS Electorate, popdensity FROM existing_and_prospective_wards_for_2023_review") %>%
  as.data.frame() %>%
  subset(select = -c(geometry))  

# Join the data together
wards <- inner_join(bc_wards, wards, by=c('ons_code' = 'ONS Code')) %>%
  as.data.frame()

# Add ONS codes and electorate to wards shapefile
shp_wards <- full_join(shp_wards, wards, by = c("NAME" = "Ward"))

# Make it a shapefile again
shp_wards <- st_sf(shp_wards, sf_column_name = 'geometry')

# Compute centroids, so we can add labels
shp_wards <- cbind(shp_wards, st_coordinates(st_centroid(shp_wards)))

# Import Polling stations
shp_polling_stations <- read_sf('./shapefiles/polling_districts.geojson')

# Define constituency colours
pal_colours <- c(
  "Calder Valley" = "Blue",
  "Halifax" = "Red",
  "Batley and Hipperholme" = "Grey"
  )

pal_shapes <- c(
  "Polling station" = 18
)

# Define plot theme
mytheme <- theme( 
  text = element_text(family = "Source Sans Pro"),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  plot.title = element_text(size = 18, face = "bold"),
  plot.subtitle = element_text(size = 12),
  axis.title = element_blank(),
  axis.line = element_blank(),
  axis.text = element_blank(),
  axis.ticks = element_blank(),
  legend.key = element_blank()
)

# Nudge labels slightly
shp_wards$nudge_y <- 0
shp_wards$nudge_x <- 0
shp_wards$nudge_y[shp_wards$abbreviation == "SB"] <- 200
shp_wards$nudge_x[shp_wards$abbreviation == "GS"] <- 650
shp_wards$nudge_x[shp_wards$abbreviation == "TOW"] <- 250
shp_wards$nudge_y[shp_wards$abbreviation == "BRI"] <- 200
```

> Within the Borough of Calderdale, the existing Halifax constituency has an electorate within the permitted range, whereas the existing Calder Valley constituency is above the permitted range.

--- [Boundary Commission Initial proposals](https://boundarycommissionforengland.independent.gov.uk/wp-content/uploads/2021/06/2021-06-08-Initial-Proposals-Yorkshire-and-the-Humber-Region.pdf)

## Constituency Boundaries

```{r, echo=FALSE, out.width="100%"}
ggplot(shp_wards)+
  mytheme+ 
  geom_sf(aes(fill = const_2019, alpha = popdensity), col = "grey20")+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )+
  scale_fill_manual(
    values = pal_colours,
    limits = names(pal_colours)[1:2], # We don't need Batley and Hipperholme for this map
  )+
  scale_alpha(range = c(0.75, 1), guide = "none")+
  geom_sf(data = shp_polling_stations, aes(shape = "Polling station"), size = 1, fill = "grey20", col = "grey20")+ # Plot polling stations on map
  scale_shape_manual(values = pal_shapes, breaks = names(pal_shapes))+
  geom_label(
    data = shp_wards, 
    size = 2,
    aes(X, Y, label = abbreviation),
    nudge_x = shp_wards$nudge_x,
    nudge_y = shp_wards$nudge_y
  )+
    guides(colour = guide_legend(order = 2), 
           shape = guide_legend(order = 1))
```

```{r, echo=FALSE}
t1 <- wards %>%
  select(const_2019, Electorate) %>%
  group_by(const_2019) %>%
  summarise(Electorate = sum(Electorate))
knitr::kable(t1, col.names = c("2019 Constituency", "Electorate"), format.args = list(big.mark = ","))
```

## (No) Simple Solutions

Evidently, there is no single ward that could be moved from Calder Valley to Halifax and still have both constituencies within the size restrictions. One or more wards would have to be exchanged, or the constituencies redrawn entirely. Wards that could be moved from Calder Valley to Halifax leaving both constituencies geographically unbroken are limited to **Hipperholme and Lightcliffe**, **Brighouse**, and **Luddenden**.

It would also be necessary to maintain some geographical consistency. It would not make sense to move **Town** ward to Calder Valley while leaving the rest of Halifax intact, for example.

**Next:** [Boundary Commission Proposal](./boundary-commission-proposal.html)