---
title: "Introduction"
author: "Tim O'Ryan"
date: "24/06/2021"
output: html_document
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(sf)
library(tidyverse)
library(data.table)
library(flextable)
library(ggspatial)

# Load shapefiles
shp_wards <- read_sf('./shapefiles/Wards_polygon.shp', query = "SELECT NAME FROM WARDS_polygon") # Calderdale ward shapefile
shp_newconst <- read_sf('./shapefiles/proposed_const.shp', 
                        query = "SELECT Constituen AS Constituency, Electorate FROM proposed_const WHERE Constituen IN (\'Halifax\', \'Calder Valley\', \'Batley and Hipperholme\')") # proposed constituencies

# Load ward data
wards <- fread('./data/calderdale_wards.csv', select=c('Ward', 'ONS Code', '2019const', 'newconst', 'abbreviation')) # 

# Load electorate from Electoral Commission ward data
bc_wards <- read_sf('./shapefiles/existing_and_prospective_wards_for_2023_review.shp', query = "SELECT ons_code, electorate AS Electorate, popdensity FROM existing_and_prospective_wards_for_2023_review") %>%
  as.data.frame() %>%
  subset(select = -c(geometry))  

# Join the data together
wards <- inner_join(bc_wards, wards, by=c('ons_code' = 'ONS Code')) %>%
  as.data.frame()

# Add ONS codes and electorate to wards shapefile
shp_wards <- full_join(shp_wards, wards, by = c("NAME" = "Ward"))

# Make it a shapefile again
shp_wards <- st_sf(shp_wards, sf_column_name = 'geometry')

# Compute centroids, so we can add labels
shp_wards <- cbind(shp_wards, st_coordinates(st_centroid(shp_wards)))

# Import Polling stations
shp_polling_stations <- read_sf('./shapefiles/polling_districts.geojson')

# Define constituency colours
pal_colours <- c(
  "Calder Valley" = "Blue",
  "Halifax" = "Red",
  "Batley and Hipperholme" = "Grey"
  )

pal_shapes <- c(
  "Polling station" = 18
)

# Define plot theme
mytheme <- theme( 
  text = element_text(family = "Source Sans Pro"),
  panel.grid = element_blank(),
  panel.background = element_blank(),
  plot.title = element_text(size = 18, face = "bold"),
  plot.subtitle = element_text(size = 12),
  axis.title = element_blank(),
  axis.line = element_blank(),
  axis.text = element_blank(),
  axis.ticks = element_blank(),
  legend.key = element_blank()
)

# Nudge labels slightly
shp_wards$nudge_y <- 0
shp_wards$nudge_x <- 0
shp_wards$nudge_y[shp_wards$abbreviation == "SB"] <- 200
shp_wards$nudge_x[shp_wards$abbreviation == "GS"] <- 650
shp_wards$nudge_x[shp_wards$abbreviation == "TOW"] <- 250
shp_wards$nudge_y[shp_wards$abbreviation == "BRI"] <- 200
```

## The Review

The Boundary Commissions for each country are holding a review of Parliamentary constituency boundaries on the basis of new rules introduced by Parliament in 2020. The rules 
require the Boundary Commissions to make more equal the number of electors in each constituency, while keeping the overall number of constituencies across the UK at 650. England has been allocated 543 constituencies for the 2023 Review, ten more than there are currently.

With the exception of two constituencies covering the Isle of Wight, each constituency must contain no fewer than 69,724 electors and no more than 77,062.

## Calderdale

```{r, echo=FALSE, out.width="100%"}
# Plot current constituencies
ggplot(shp_wards)+
  mytheme+ 
  geom_sf(aes(alpha = popdensity), col = "grey20")+
  theme(
    legend.position = "bottom",
    legend.title = element_blank()
  )+
  scale_alpha(range = c(0.75, 1), guide = "none")+
  geom_sf(data = shp_polling_stations, aes(shape = "Polling station"), size = 1, fill = "grey20", col = "grey20")+ # Plot polling stations on map
  scale_shape_manual(values = pal_shapes, breaks = names(pal_shapes))+
  geom_label(
    data = shp_wards, 
    size = 2,
    aes(X, Y, label = abbreviation),
    nudge_x = shp_wards$nudge_x,
    nudge_y = shp_wards$nudge_y
  )
```

```{r echo=FALSE}
t1 <- subset(wards, select=c("Ward", "Electorate", "abbreviation"))
func <- function(z) if (is.numeric(z)) sum(z) else ''
sumrow <- as.data.frame(lapply(t1, func))
sumrow[1] <- "**Total**"
knitr::kable(rbind(t1, sumrow),
             col.names = c("Ward", "Electorate*", "Abbreviation"), 
             caption = "* Boundary Commission for England initial proposal, June 2021", 
             format.args = list(big.mark = ",")) # 
```

**Next:** [The current position](./current-position.html)
